### 2022-10-04 chapters/optimization/java-dir/java.tex
Proofreading conclusions

\clearpage
\section{Summary}

This chapter describes an empirical investigation of contrast in energy consumption that some of the most well-known and widely supported JVM platforms can exhibit as well as the important factors that can influence this energy consumption.
During our experiments, we investigated 12 well-known and diverse-purpose Java benchmarks, as well as 52 JVMs, including many versions of 11 different distributions.
Our findings revealed that various JVMs share energy efficiency and may be classified into three major classes: \textsc{HotSpot}, \textsc{J9}, and \textsc{GraalVM}.
However, the three JVM classes can show substantially different energy efficiency depending on the workloads and/or application.
While there was no clear champion in terms of energy consumption,\textsc{GraalVM} recorded the highest energy efficiency for the most of benchmarks. Nevertheless, depending on the workload, each JVM can achieve higher or worse efficiency.
One explanation could be thread management policy, as seen with \textsc{J9} when running \textsf{Avrora}.

@@

On the other hand , the default GC outperformed the other strategies in just half of our experiments, with some significant benefits reported when employing different GCs based on the application characteristics. Our primary observations and recommendations can be stated as follows:
\begin{inparaenum}[\em i)]
    \item testing software on the 3 classes of JVM and identifying the one that consumes the least is a good practice, especially for multi-threading purposes,
    \item while the JVM default JIT give often good energy consumption results, some settings may improve the energy consumption and could be tested,
    \item the choice of the GC may lead to a large impact on the energy consumption in many situations, thus encouraging a careful tuning of this parameter before deployment.
\end{inparaenum}

To make the above principles easier to implement, we propose \emph{J-Referral}, a tool that recommends the best energy-efficient JVM distribution and configuration from among more than a hundred options.
It generates a comprehensive report on the energy consumption of both CPU and DRAM components for each JVM distribution and/or configuration in order to assist the user in selecting the one with the lowest consumption for Java applications.

>>>>

\clearpage
\section{conclusion}
This chapter describes an empirical investigation of contrast in energy consumption that some of the most well-known and widely supported JVM platforms can exhibit as well as the important factors that can influence this energy consumption.
During our experiments, we investigated 12 well-known and diverse-purpose Java benchmarks, as well as 52 JVMs, including many versions of 11 different distributions.
Our findings revealed that various JVMs share energy efficiency and may be classified into three major classes: \textsc{HotSpot}, \textsc{J9}, and \textsc{GraalVM}.
However, the three JVM classes can show substantially different energy efficiency depending on the workloads and/or application.
While there was no clear champion in terms of energy consumption,\textsc{GraalVM} recorded the highest energy efficiency for the most of benchmarks. Nevertheless, depending on the workload, each JVM can achieve higher or worse efficiency.
One explanation could be thread management policy, as seen with \textsc{J9} when running \textsf{Avrora}.

@@

On the other hand, the default GC outperformed the other strategies in just half of our experiments, with some significant benefits reported when employing different GCs based on the application characteristics. Our primary observations and recommendations can be stated as follows:
\begin{inparaenum}[\em i)]
    \item testing software on the 3 classes of JVM and identifying the one that consumes the least is a good practice, especially for multi-threading purposes,
    \item while the JVM default JIT give often good energy consumption results, some settings may improve the energy consumption and could be tested,
    \item the choice of the GC may lead to a large impact on the energy consumption in many situations, thus encouraging a careful tuning of this parameter before deployment.
\end{inparaenum}

To make the above principles easier to implement, we propose \emph{J-Referral}, a tool that recommends the best energy-efficient JVM distribution and configuration from among more than a hundred options.
It generates a comprehensive report on the energy consumption of both CPU and DRAM components for each JVM distribution and/or configuration to assist the user in selecting the one with the lowest consumption for Java applications.

### 2022-10-04 chapters/programming_languages/programming_languages.tex
remove : link
Proofreading conclusions

\section{Investigating Web Application Frameworks}\label{sec:webframework}
\import{\currfiledir/web_frameworks}{web_frameworks}

\section{Conclusion \note{missing}}

>>>>

\section{Investigating Web Application Frameworks}\label{sec:webframework}

\import{\currfiledir/web_frameworks}{web_frameworks}

\section{Conclusion}
This chapter aims to extend the work of \citeauthor{pereira_energy_2017}~\cite{pereira_energy_2017} by studying the energy footprint on other aspects than micro-benchmarking.
The main focus of this chapter is the behavior of programming languages within the context of web applications.
To do so, we have covered two case studies.
The first experiment is the impact of programming languages when developing RPC services.
We used the official implementation of \textsc{gRPC} library to build $25$ frameworks using $17$ programming languages.
In the second experiment, we look at how web frameworks affect energy consumption by creating a basic web application and comparing its performance, latency, and energy usage using the most common web frameworks.
Overall we had $261$ implementation using $35$ and three different databases.
For each implementation, we considered $7$ representative scenarios.
Each scenario is tested with several stress levels to study the scalability of the web frameworks.
Therefore, the final results cover $8,750$ use cases.

Our studies revealed the absence of a clear winner in terms of energy efficiency.
However, when we take into consideration the cost of a single request.
C++ and Rust were the top-ranked, hence assessing the work done by \citeauthor{pereira_energy_2017}.
On the other hand, Unlike the results shown in the paper \cite{pereira_energy_2017}, \textsf{PHP} exhibits a higher performance while keeping a high power consumption on average.
This result is due to the nature of \textsf{PHP} language, which is dedicated to the web and not general-purpose programming. 
While most of the frameworks consumed around $100 watts$, \textsf{Java}-based frameworks consumed around $220 watts$.
This made \textsf{Java} one of the most energy-consuming in the context of online services, especially when considering the periods with low traffic.

Finally, this study revealed that each framework had different behaviors towards scalability.
Some of them increased the average power consumption while trying to keep up with the performance.
Others, on the other hand, could keep the average power consumption while decreasing the performance.
Furthermore, there was a case where there was a mix between both strategies.

As there is no final answer to the question of which framework is the most energy-efficient, we provide developers with a tool to help them compare between frameworks based on their needs.
The criteria we used to compare frameworks are the scenario type, the stress level, and the database type.

% The tool is available at \url{https://github.com/chakib-belgaid/greenboard}

### 2022-10-04 chapters/programming_languages/rpc/rpc.tex
add : summary for programmming langaugess
Proofreading conclusions

\subsection{Conclusion}

This section presents a study on the impact of programming languages on the energy for request handling. To do that, we used an implementation of 25 frameworks based on 17 programming languages.
These servers use the official RPC protocol, reducing programmers' bias toward specific languages.

The results show that the number of clients had a binary effect on the servers.
With a small number of clients, there was no typical behavior between different categories. On the other hand, when dealing with a more significant number of clients,  interpreted languages and compiled ones dropped in terms of performance while keeping the same average power, unlike the Vm~based ones that increased their average power while keeping the same performances. Overall the second strategy had a better impact on the single request cost than the first one.

As for the evolution of the query size, we have seen three modes of behavior for each framework, depending on the size of the requests: the stress-free mode, the escalation mode, and the broken state mode.
We start with the stress-free mode. The leader class was the compiled languages, while the VM-based ones tended to consume more energy.
Finally, with the escalation mode. one could observe three strategies to deal with the increase in request size. The first strategy was to drop in performance while keeping the same power; the second was to increase the average power consumption without impacting the performance; the third was a combination of both previous strategies with an increase in power and a drop in performance. Interestingly the third strategy had a lower cost per request than the first one.
Finally, most implementations could not handle requests heavier than 10 MB for the broken mode. Elixir was an exception to this limit despite its low performance compared to other languages

>>>>

\subsection{Summary}

This section reports on a study on the impact of programming languages on the energy for request handling.
To do that, we used an implementation of $25$ frameworks based on $17$ programming languages.
These servers use the official RPC protocol, reducing programmers' bias toward specific languages.

The results show that the number of clients had a binary effect on the servers.
With a small number of clients, there was no typical behavior between different categories.
On the other hand, when dealing with a more significant number of clients, interpreted languages and compiled ones dropped in terms of performance while keeping the same average power, unlike the VM-based ones that increased their average power while keeping the same performances.
Overall, the second strategy had a better impact on the single request cost than the first one.

As for the evolution of the query size, we have seen three modes of behavior for each framework, depending on the size of the requests: the stress-free mode, the escalation mode, and the broken state mode.
We start with the stress-free mode.
The leader class was the compiled languages, while the VM-based ones tended to consume more energy.
Finally, with the escalation mode, one could observe three strategies to deal with the increase in request size.
The first strategy tends to drop performances while consuming the same power; the second was to increase the average power consumption without impacting the performances; the third was a combination of both previous strategies with an increase in power and a drop in performance.
Interestingly, the third strategy had a lower cost per request than the first one.
Finally, most implementations could not handle requests heavier than 10 MB for the broken mode.
Elixir was an exception to this limit despite its low performance compared to other languages.
